### 2022-08-23 main/chapter_searches.tex
mostly adding citations; some content on binomial

\paragraph{Probability assignments.}
Probability algebra manipulates probabilities, but does not state how they
should be initially assigned.
Probability assignments can, in general, be freely assigned and refined from
calibration data. But there are unambiguous cases that are commonly used in
High-Energy, notably the Poisson and Gaussian distributions.
To avoid turning this thesis into a textbook, we only sketch the derivations
of these, and refer to real textbooks for their detailed
derivations~\cite{jaynes2003probability}.

Both Poisson and
Binary accept/reject experiments are described by the
The Large Hadron Collider creates $40$~million proton collisions per second,
only around $1000$ are accepted by \atlas' trigger system per
second~\cite{atlas2020trigger}.
This acceptance rate of $250$ parts per million is sufficient to invoke the
Poisson distribution for event counts.
Each collision event has some probability $p$ of passing the trigger
\begin{equation}
\label{eqn:searches_poisson}
\mathrm{Poisson}(n\mid \lambda) = \frac{\lambda^n}{n!}\exp{(-\lambda)}
\end{equation}

Succinctly stated by Jaynes that
``any function with a single rounded maximum, raised to a higher and higher
power, goes into a Gaussian function''~\cite{jaynes2003probability}.
\begin{equation}
\label{eqn:searches_gaussian}
\mathrm{Gaussian}(x\mid \mu, \sigma) =
\frac{1}{\sqrt{2\pi \sigma^2}}
\exp{\left(-\frac{1}{2}\frac{(x - \mu)^2}{\sigma^2}\right)}
\end{equation}
This is not the famous Central Limit Theorem, which is related to the
distribution of sums, but can be used to derive it: the distribution of a sum
is a convolution integral, which becomes a product in Fourier-transformed
space, and that product can converge to a Gaussian function by this rule.
Jaynes gives a beautiful alternative proof of the
Central Limit Theorem~\cite{jaynes2003probability}
(Section ``7.16 The central limit theorem'').

>>>>

\paragraph{Probability assignments.}
Probability algebra manipulates probabilities, but does not state how they
should be initially assigned.
In general, such probability assignments are free and unconstrained, and might
informed by calibration data or loosely motivated opinions.
But there are examples of unambiguously derived that are commonly used in
High-Energy Physics, notably the Poisson and Gaussian distributions.
To avoid turning this thesis into a textbook, we only sketch the derivations
of these, and refer to real textbooks for their detailed
derivations~\cite{jaynes2003probability}, both of which use the exponential
identity that
\begin{equation}
\label{eqn:searches_exponential}
\exp(x) = \displaystyle \lim_{N \to \infty} (1 - \frac{x}{N})^N
.
\end{equation}

The Large Hadron Collider collides $40$~million proton bunches per second,
but \atlas' trigger system accepts only only around $1000$ of
those~\cite{atlas2020trigger}.
This is a Binomial scenario with a large fixed number $N$ of trials
(bunch collisions)
and a small probability of acceptance (triggering),
due to a notable collision event.
The Binomial distribution for $n$ triggered events is well known as
\begin{equation}
\mathrm{Binomial}(n\mid N, p) = p^n (1 - p)^{N - n} \times \binom{N}{n}
,
\end{equation}
which is a product of the probabilities to the sequence of $N$ accept/reject
data, summed over the permutations that could have led to the total $n$.
With our extremely small trigger rate, we are interested in the limit of small
$p$ and large $N$ which result in a mean acceptance number of $\lambda = Np$.
In the large $N$ limit, this tends towards the Poisson
distribution~\cite{jaynes2003probability}
\begin{equation}
\label{eqn:searches_poisson}
\mathrm{Poisson}(n\mid \lambda) = \frac{\lambda^n}{n!}\exp(-\lambda)
,
\end{equation}
which is used ubiquitously in the analysis of our histograms.

The Gaussian distribution has many justifications:
as a practical approximation it permits analysis in linear algebra,
as the unique distribution is uniquely spherically symmetric and factors into
independent dimensions~\cite{
jaynes2003probability,
herschel1850normal,
maxwell1860normal,
muller1959note,
marsaglia1972normal
},
and as the maximum entropy~\cite{PhysRev.106.620} distribution for a
known mean and variance.
\TODO{}
Succinctly stated by Jaynes that
``any function with a single rounded maximum, raised to a higher and higher
power, goes into a Gaussian function''~\cite{jaynes2003probability}.
\begin{equation}
\label{eqn:searches_gaussian}
\mathrm{Gaussian}(x\mid \mu, \sigma) =
\frac{1}{\sqrt{2\pi \sigma^2}}
\exp\left(-\frac{1}{2}\frac{(x - \mu)^2}{\sigma^2}\right)
\end{equation}
This is not the famous Central Limit Theorem, which is related to the
distribution of sums, but can be used to derive it: the distribution of a sum
is a convolution integral, which becomes a product in Fourier-transformed
space, and that product can converge to a Gaussian function by this rule.
Jaynes gives a beautiful alternative proof of the
Central Limit Theorem~\cite{jaynes2003probability}
(Section ``7.16 The central limit theorem'').
